Investigating social change during cultural contact period using geometric morphometry of pottery shapes from Iron Age northeastern Taiwan

### 2020-01-26 analysis/paper/paper.Rmd
para 1

The development of craft specialization is usually associated with the social changes since it reflects production systems that could inform social, economic, or political conditions of a society. Specialization is defined as "differential participation in specific economic activities", in which a particular good is produced by fewer producers compared to the total population who consume it [@Costin1991; @Costin2001, pp. 43]. The implication of the emergence of specialization for social structure or social process has been a theme in archaeological investigations. From the perspective of the political approaches, craft specialization specifically attached production, is a strategy used by elites to control over production systems to maintain power and authority [@Costin2001]. The relationship between craft specialization and social differentiation is broadly discussed, and many cases show that specialized production is a useful indicator of increasing socio-political complexity [@Hirshman2010; @Junker1999]. One the other hand, economic models consider production, exchange, and consumption as a whole to explain specialization as a subsistence strategy to accommodate uneven distribute of resources [@Costin2001; @Arnold1994], a reaction to the changes in subsistence practices [@Stark1995economic], or an outcome result from by long-distance trade and inter-regional exchange [@Alizadeh2018]. An present-day investigation of ceramic specialization in the Jodhpur region of India suggests that the adoption of standardized products across cultural boundaries is a reaction to a collapse of previous economic system, indicating the correlation with major socio-economic changes [@Roux2015]. However, it is important to note that the explanation for the emergence of specialized production might not be straightforward or exclusive. @Acabado2018's case study in lowland northern Philippines shows that specialization is not necessarily accompanied with social differentiation based on the result of a low degree of ceramic specialization in Ifugao ranked societies during the Spanish colonization. The development of specialization could be history dependent, where condition and contexts under which specialists organized are important to make an inference for changes in social organization [@Roux2015; @Costin2001].

>>>>

The development of craft specialization is sometimes associated with social changes since it reflects changes in production systems that could inform on the social, economic, or political conditions of a society. We define specialization as "differential participation in specific economic activities", in which a particular good is produced by fewer producers compared to the total population who consume it [@Costin1991; @Costin2001, pp. 43]. The effects of the emergence of specialization in past societies on their social structures and social processes is an important question in archaeological investigations of small scale societies in the Late Holocene. Considered through a political lens, craft specialization may be a strategy used by elites to control production systems to maintain their power and authority [@Costin2001]. This relationship between craft specialization and social differentiation has been broadly discussed, and many cases show that increasingly specialized production can be a useful indicator of increasing socio-political complexity [@Hirshman2010; @Junker1999]. That said, more economic focus considers production, exchange, and consumption as a whole to explain specialization as part of a strategy to accommodate uneven distributions of resources [@Costin2001; @Arnold1994]. Craft specialisation may appear as a reaction to changes in subsistence practices [@Stark1995economic], or as a result of long-distance trade and inter-regional exchange [@Alizadeh2018]. Present-day investigation of ceramic specialization in the Jodhpur region of India suggests that the adoption of standardized products across cultural boundaries is a reaction to a collapse of previous economic system, indicating a correlation with major socio-economic changes [@Roux2015]. However, it is important to note that explanations for the emergence of specialized production in past societies might not be so straightforward or singular. @Acabado2018's case study in lowland northern Philippines shows that specialization is not necessarily accompanied with social differentiation based on a low degree of ceramic specialization in Ifugao ranked societies during the time of Spanish colonization. The development of specialization could be dependent on unique historical factors, where the conditions and contexts under which specialists organized must be taken into account when making inferences about changes in social organization [@Roux2015; @Costin2001].
